History of Social Work Research: Data Cleaning

### 2015-02-03 dataCleaning.Rmd
More data cleaning.

This document summarizes data cleaning procedures of the data file processed by the `ebscoBWR` fucntion.  It should be noted that that `ebscoBWR` function merged the search results from the first- and second-level searches performed on the three databases (psycInfo, Social Science Abstracts, and Social Work Abstract) hosted by the EbscoPlatform.  This function merged all the search results, transformed the data into a structured data file, and removed all duplicate article records.  No further transformations or cleaning was performed by the `ebscoBWR` function.  Provided below are basic descriptive summaries of the results from the full search before further cleaning.

It should be noted that the second-level search was specified to be overly inclusive to help ensure the most comprehensive search possible.  All journal titles from the secondary search are identified in the following summary.

library(dplyr)
temp <- bwr.df[bwr.df$articleID == 28300,]
27743 and 28300.

@@

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
 # Clear workspace
rm(list=ls())

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/ebscoBWR.R")

# Set the path where original raw data are stored
setwd("/Users/beperron/Git/SocialWorkResearch/Data")

# Set the working directory to store files created by BWR functions
path <- "/Users/beperron/Git/SocialWorkResearch/Data"
```

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("/Users/beperron/Git/SocialWorkResearch/Data/HistoricalArticles.R")

library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)
library(reshape2)
```

# Data Pre-Processing

# NOTE: CONFIRM THAT SOCIAL CASEWORK IS SAME AS FAMILIES IN SOCIETY.  ENSURE WE 
# HAVE THE ARTICLES FOR WHEN THE JOURNAL WAS ALSO CALLED THE FAMILY
# Re-write this to call the R-object directly
# Overview of processed datafile

Provided below are some basic descriptive statistics before performing any data cleaning.

```{r echo=FALSE, comment=NA} 
number.of.articles <- length(which(bwr.df$attributes == "TI"))
range.of.years <- range(bwr.df$record[bwr.df$attributes == "YR"])
print(number.of.articles)
print(range.of.years)
```

```{r echo=FALSE, comment=}
unique.titles <- filter(bwr.df, attributes == "SO") %>%
    group_by(record) %>%
    summarise(
        N = n())
print.data.frame(unique.titles, rowNames=FALSE)
```

@@

```{r}
years <- c("2014", "2015")
ebsco.remove <- bwr.df[bwr.df$record %in% years, ]
ebsco.remove <- ebsco.remove[, "articleID"]
bwr.df <- bwr.df[!(bwr.df$articleID %in% ebsco.remove),]
```

@@

```{r}
bwr.df$record <- ifelse(bwr.df$record == "Eugenics and Social Welfare Bulletin", 
                       "Eugenics & Social Welfare Bull.", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "The British Journal of Social Work", 
                       "British Journal of Social Work", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Social Work & Social Sciences Review", 
                       "Social Work and Social Sciences Review", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Social Work with Groups", 
                       "Social Work With Groups", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Computers in Human Services", 
                       "Journal of Technology in Human Services", 
                        bwr.df$record)

bwr.df$record <- ifelse(bwr.df$record == "Journal of Sociology & Social Welfare", 
                       "Journal of Sociology and Social Welfare", 
                        bwr.df$record)

#Note leave space after the (Electronic only)
bwr.df$record <- ifelse(bwr.df$record == "Contemporary Rural Social Work (Electronic only) ", 
                        "Contemporary Rural Social Work (Electronic only) ", bwr.df$record)

@@

```{r}
exclusions <- c(
    "American Journal of Preventive Medicine", 
    "Behavior Modification", 
    "Brain and Cognition",
    "Canadian Journal of Community Mental Health",
    "Canadian Journal on Aging",
    "Early Child Development and Care",
    "Early Education and Development",
    "Employee Assistance Quarterly",
    "Evaluation and Program Planning",
    "General Hospital Psychiatry",
    "Hospital Social Service",
    "International Review of Social Research",
    "Journal of Applied Behavioral Science",
    "Journal of Applied Rehabilitation Counseling",
    "Journal of Community & Applied Social Psychology",
    "Journal of Health & Human Services Administration",
    "Journal of Nonverbal Behavior",
    "Learning Disabilities Research & Practice",
    "New York State Department of Social Welfare, Division Publication",
    "Pennsylvania Social Work",
    "Proceedings of the National Converence of Social Work",
    "PROFILE Issues in Teachers' Professional Development",
    "Psychotherapy",
    "Social Service Department, Presbyterian Hospital, New York",
    "Social Work Technique",
    "Social Work Today", # Magazine
    "Social Work Yearbook",
    "The New Social Worker",
    "Systems Research and Behavioral Science",
    "Teaching and Teacher Education",
    "University of Chicago Social Service Monograph",
    "The Scientific Review of Mental Health Practice",
    "The Clinical Supervisor",
    "University of Chicago Social Service Monograph")
ebsco.remove <- bwr.df[bwr.df$record %in% exclusions, ]
ebsco.remove <- ebsco.remove[, "articleID"]
bwr.df <- bwr.df[!(bwr.df$articleID %in% ebsco.remove),]
```

```{r unique_titles, echo=FALSE}
unique.titles <- filter(bwr.df, attributes == "SO") %>%
    group_by(record) %>%
    summarise(
        N = n())
```

@@

```{r unique_titles, echo=FALSE}
unique.titles <- filter(bwr.df, attributes == "SO") %>%
    group_by(record) %>%
    summarise(
        N = n())

title <- bwr.df %>%
        filter(attributes == "SO") %>%
        select(id = articleID, SO = record)

article <- bwr.df %>%
        filter(attributes == "TI") %>%
        select(id = articleID, journal.title = record)

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(bwr.df)

@@

n.so.yr <- filter(bwr.df, attributes == "SO" | attributes == "YR")

n.so <- filter(bwr.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(bwr.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

@@

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(bwr.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(bwr.df, attributes == "KP")

@@

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(bwr.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- bwr.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(bwr.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

@@

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}
library(RColorBrewer)
year <- filter(bwr.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

location.temp <- bwr.df %>%
        filter(attributes == "LO") %>%
        select(articleID = articleID, country = record) %>%
        mutate(country = gsub(" ", "", country))

@@

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- bwr.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(bwr.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

>>>>

This document summarizes data cleaning procedures of the data file processed by the `ebscoBWR` fucntion.  It should be noted that that `ebscoBWR` function merged the search results from the first- and second-level searches performed on the three databases (psycInfo, Social Science Abstracts, and Social Work Abstract) hosted by the EbscoPlatform.  This function merged all the search results, transformed the data into a structured data file, and removed all duplicate article records.  No further transformations or cleaning was performed by the `ebscoBWR` function.  Provided below are basic descriptive summaries of the results from the full search before further cleaning.  It should be noted that the second-level search was specified to be overly inclusive to help ensure the most comprehensive search possible.  All journal titles from the secondary search are identified in the following summary.

The main file for analysis is `fullDF.R`, and a separate file is just for psycInfo `psycInfoDF.R`.  The same cleaning procedures need to be performed on both data files prior to analysis.

# Initialization of workspace

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=TRUE}
# Clear workspace
rm(list=ls())

# Load required packages
library(ggplot2)
library(dplyr)

# Read data files
load("/Users/beperron/Git/SocialWorkResearch/Data/psycInfoDF.R")
load("/Users/beperron/Git/SocialWorkResearch/Data/fullDF.R")

# Set the working directory
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
```

# Pre-processing summary

Provided below are some basic descriptive statistics before any data of the files processed by `ebscoBWR` function.

```{r echo=FALSE, comment=NA} 
number.of.articles <- length(which(full.df$attributes == "article"))
range.of.years <- range(full.df$record[full.df$attributes == "pubYear"])
cat(number.of.articles)
cat(range.of.years)
```

*TO DO* add range of years to the following dplyr summary

```{r echo=FALSE, comment=}
unique.titles <- filter(full.df, attributes == "journal") %>%
    group_by(record) %>%
    summarise(
        N = n())
print.data.frame(unique.titles, rowNames=FALSE)
```

# Data cleaning

The first step is to exclude years 2014 and 2015 due to delays in indexing.

```{r}
years <- c("2014", "2015")
ebsco.remove <- full.df[full.df$record %in% years, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]
```

The articles were reviewed and a first round of exclusions were made.

```{r}
exclusions <- c(
    "American Journal of Preventive Medicine", 
    "Behavior Modification", 
    "Brain and Cognition",
    "Canadian Journal of Community Mental Health",
    "Canadian Journal on Aging",
    "Early Child Development and Care",
    "Early Education and Development",
    "Employee Assistance Quarterly",
    "Evaluation and Program Planning",
    "General Hospital Psychiatry",
    "Hospital Social Service",
    "International Review of Social Research",
    "Journal of Applied Behavioral Science",
    "Journal of Applied Rehabilitation Counseling",
    "Journal of Community & Applied Social Psychology",
    "Journal of Health & Human Services Administration",
    "Journal of Nonverbal Behavior",
    "Learning Disabilities Research & Practice",
    "New York State Department of Social Welfare, Division Publication",
    "Pennsylvania Social Work",
    "Proceedings of the National Converence of Social Work",
    "PROFILE Issues in Teachers' Professional Development",
    "Psychotherapy",
    "Social Service Department, Presbyterian Hospital, New York",
    "Social Work Technique",
    "Social Work Today", # Magazine
    "Social Work Yearbook",
    "The New Social Worker",
    "Systems Research and Behavioral Science",
    "Teaching and Teacher Education",
    "University of Chicago Social Service Monograph",
    "The Scientific Review of Mental Health Practice",
    "The Clinical Supervisor",
    "University of Chicago Social Service Monograph")
ebsco.remove <- full.df[full.df$record %in% exclusions, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]

rm(ebsco.remove)
```

Journal titles were merged in the following situations:

1.
2.
3.

```{r}
full.df$record <- ifelse(full.df$record == "Eugenics and Social Welfare Bulletin", 
                       "Eugenics & Social Welfare Bull.", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "The British Journal of Social Work", 
                       "British Journal of Social Work", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Social Work & Social Sciences Review", 
                       "Social Work and Social Sciences Review", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Social Work with Groups", 
                       "Social Work With Groups", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Computers in Human Services", 
                       "Journal of Technology in Human Services", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Journal of Sociology & Social Welfare", 
                       "Journal of Sociology and Social Welfare", 
                        full.df$record)

#Note leave space after the (Electronic only)
full.df$record <- ifelse(full.df$record == "Contemporary Rural Social Work (Electronic only) ", 
                        "Contemporary Rural Social Work (Electronic only) ", full.df$record)

@@

@@

```{r}
years <- c("2014", "2015")
ebsco.remove <- full.df[full.df$record %in% years, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]
```

@@

```{r}
full.df$record <- ifelse(full.df$record == "Eugenics and Social Welfare Bulletin", 
                       "Eugenics & Social Welfare Bull.", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "The British Journal of Social Work", 
                       "British Journal of Social Work", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Social Work & Social Sciences Review", 
                       "Social Work and Social Sciences Review", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Social Work with Groups", 
                       "Social Work With Groups", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Computers in Human Services", 
                       "Journal of Technology in Human Services", 
                        full.df$record)

full.df$record <- ifelse(full.df$record == "Journal of Sociology & Social Welfare", 
                       "Journal of Sociology and Social Welfare", 
                        full.df$record)

#Note leave space after the (Electronic only)
full.df$record <- ifelse(full.df$record == "Contemporary Rural Social Work (Electronic only) ", 
                        "Contemporary Rural Social Work (Electronic only) ", full.df$record)

@@

```{r}
exclusions <- c(
    "American Journal of Preventive Medicine", 
    "Behavior Modification", 
    "Brain and Cognition",
    "Canadian Journal of Community Mental Health",
    "Canadian Journal on Aging",
    "Early Child Development and Care",
    "Early Education and Development",
    "Employee Assistance Quarterly",
    "Evaluation and Program Planning",
    "General Hospital Psychiatry",
    "Hospital Social Service",
    "International Review of Social Research",
    "Journal of Applied Behavioral Science",
    "Journal of Applied Rehabilitation Counseling",
    "Journal of Community & Applied Social Psychology",
    "Journal of Health & Human Services Administration",
    "Journal of Nonverbal Behavior",
    "Learning Disabilities Research & Practice",
    "New York State Department of Social Welfare, Division Publication",
    "Pennsylvania Social Work",
    "Proceedings of the National Converence of Social Work",
    "PROFILE Issues in Teachers' Professional Development",
    "Psychotherapy",
    "Social Service Department, Presbyterian Hospital, New York",
    "Social Work Technique",
    "Social Work Today", # Magazine
    "Social Work Yearbook",
    "The New Social Worker",
    "Systems Research and Behavioral Science",
    "Teaching and Teacher Education",
    "University of Chicago Social Service Monograph",
    "The Scientific Review of Mental Health Practice",
    "The Clinical Supervisor",
    "University of Chicago Social Service Monograph")
ebsco.remove <- full.df[full.df$record %in% exclusions, ]
ebsco.remove <- ebsco.remove[, "articleID"]
full.df <- full.df[!(full.df$articleID %in% ebsco.remove),]
```

```{r unique_titles, echo=FALSE}
unique.titles <- filter(full.df, attributes == "SO") %>%
    group_by(record) %>%
    summarise(
        N = n())
```

@@

```{r unique_titles, echo=FALSE}
unique.titles <- filter(full.df, attributes == "SO") %>%
    group_by(record) %>%
    summarise(
        N = n())

title <- full.df %>%
        filter(attributes == "SO") %>%
        select(id = articleID, SO = record)

article <- full.df %>%
        filter(attributes == "TI") %>%
        select(id = articleID, journal.title = record)

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(full.df)

@@

n.so.yr <- filter(full.df, attributes == "SO" | attributes == "YR")

n.so <- filter(full.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(full.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

@@

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(full.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(full.df, attributes == "KP")

@@

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(full.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- full.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(full.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

@@

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}
library(RColorBrewer)
year <- filter(full.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

location.temp <- full.df %>%
        filter(attributes == "LO") %>%
        select(articleID = articleID, country = record) %>%
        mutate(country = gsub(" ", "", country))

@@

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- full.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(full.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)
